Solving non smooth dynamics using ADMM

### 2020-10-21 2021/ADMM_journal/author.tex
Add stub of article for a journal, on ADMM topic - better and more detailed version of the Nodycon2021 paper

%%%%%%%%%%%%%%%% based on Springer template for conference proceedings %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{svproc}

% to typeset URLs, URIs, and DOIs
\usepackage{url}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{csquotes}

\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{cleveref}
\usepackage{bm}
\usepackage{mathtools}
\usepackage{siunitx}
\usepackage{subfig}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{tabularx}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{array}
\usepackage{booktabs}
\usepackage{multicol}
\usepackage{xcolor}
\usepackage[title]{appendix}
\usepackage{empheq}

@@

@@

\begin{document}
\mainmatter              % start of a contribution
%
\title{Solving Non-smooth Dynamic Problems using \\ the 
Alternating Direction Method of Multipliers}
%
\titlerunning{Non-smooth dynamic problem}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Alessandro Tasora \and Dario Mangoni \and Simone Benatti}
%
\authorrunning{Alessandro Tasora et al.} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\tocauthor{Alessandro Tasora, Dario Mangoni, and Simone Benatti}
%
\institute{University of Parma, Department of Engineering and Architecture \\ V.le delle Scienze 181/A, 40121 ITALY\\
\email{alessandro.tasora@unipr.it},\\ WWW home page:
\texttt{http://digitaldynamicslab.unipr.it}
}

@@

\begin{abstract}
We propose to use the Alternating Direction Method of Multipliers (ADMM) for solving the variational inequality that arises in non-smooth contact problems in multibody simulations. The ADMM method is based on simple computational primitives and offers good convergence properties especially for scenarios where loose tolerance on the precision can be accepted, thus making it attractive also for real-time applications. The method scales well with the problem size and require a small number of tuning parameters. Differently from other methods known in literature, such as fixed point iterations, it can easily handle problems that mix finite elements and rigid bodies, and it is only marginally affected by odd mass ratios or ill-posed problems. Another attractive feature is that it can be easily warm-started.
\keywords{ADMM, non-smooth dynamics, contact}
\end{abstract}
%

@@

Recently there has been a revival of Alternating Direction Method of Multipliers (ADMM) and similar operator-splitting methods for solving constrained optimization problems
\cite{Goldstein2014}. %  [2014 Alternating Direction Optimization Methods]. 
Even if the theoretical convergence of such methods is worse than the one of IPMs, recent developments showed that in practical scenarios they offer superior speed, scalability and robustness 
\cite{Cannon2019}. % [COSMO  conic operator splitting method for large convex problems].
An attractive property of ADMM is that it requires the solution of a linear system per each iteration, but unlike IPMs, such linear system most often remains unchanged during the iterations, so a factorization can be reused multiple times with a great benefit in terms of speed.

In 
\cite{Stellato2020} % [OSQP: an operator splitting solver for quadratic programs].
a specialized ADMM method for quadratic problems with conic constraints has been presented: on the top of this method we develop a solver that can exploit the nature of non-smooth dynamical problems.

% TO ADD IN EXTENDED VERSION
%In our embodiment, unknowns are primal variables (velocity measures) and dual variables (impulses in contact points and joints), cone constraints stem from the Coulomb friction law, and the system matrix includes terms from the mass matrices and from the tangent stiffness/damping matrices (hence it is sparse, but not necessarily diagonal).
%This paper will present a model for the non-smooth dynamics, it will show how to cast it as a QP with convex conic constraint, it will present the ADMM method to solve it, it will discuss practical implementation details and then it will show some benchmarks.

@@

We remark that the ADMM method just makes the assumption of $\Upsilon$ being convex, so the problem can be generalized to $\vect{\gamma}_\epsilon \in \mathcal{C}$ where $\mathcal{C}$ is a generic convex set and at the same time we assume an associated 
% TO ADD IN EXTENDED VERSION
%\footnote{Not all dissipative constitutive models are associated, for example in some computational plasticity models the plastic flow might deviate respect to the normal to the yeld surface, and the Coulmb contact model itself is not associated -in fact the contact velocity $\vect{u}$ can range into a wider cone than the polar of the friction cone, but our DVI model compensates fur such effect. We will assume associated models heretofore, and we do not deal with non-associated models to avoid additional complexity for the moment.} 
flow $\vect{u}_\epsilon \in - \mathcal{N}_{C}(\vect{\gamma}_\epsilon)$. For instance, $\mathcal{C}$ could be a capped friction cone to represent plasticization of contacts, or a cone translated downward to represent possible adhesion in contact up to a threshold, or the Von Mises yield region if $\vect{\gamma}_\epsilon$ represent stresses in finite elments undergoing plasticization.

@@

For variables $(\vect{x},\vect{z}) \in \mathbb{R}^m \times \mathbb{R}^n$, the ADMM methods solve constrained optimization problems with separable structure
\begin{subequations}
	\begin{align}
    \text{min} \quad  & f(\vect{x}) + g(\vect{z}) \\
	  \text{s.t.} \quad &  A \vect{x} + B \vect{z} - \vect{b} = \vect{0}
	\end{align}
	\label{eq:admm}
\end{subequations}
by introducing an augmented Lagrangian
\begin{align}
\mathcal{L}_{\rho}(\vect{x},\vect{z},\vect{y})= f(\vect{x}) + g(\vect{z}) + \vect{y}^T (A \vect{x} + B \vect{z} - \vect{b}) + \frac{\rho}{2} \norm{A \vect{x} + B \vect{z} - \vect{b}}
\end{align}
and by iterating over two minimization problems as in the following loop:
\begin{align}
 \vect{x}^{k+1} &\in \text{argmin} \mathcal{L}_{\rho}(\vect{x},\vect{z}^k,\vect{y}^k) \\
 \vect{z}^{k+1} &\in \text{argmin} \mathcal{L}_{\rho}(\vect{x}^{k+1},\vect{z},\vect{y}^k) \\
 \vect{y}^{k+1} &= \vect{y}^{k} + \rho ( A \vect{x}^{k+1} + B \vect{z}^{k+1} - \vect{b} )
 \label{eq:admm_iters}
\end{align}
Although the convergence is fast in the first iterations, it tends to deteriorate later as in fixed-point iterations; however in practical scenarios where loose tolerances can be accepted, the method proves to be very efficient and robust.

@@

We can write the augmented Lagrangian introducing two step size parameters $\sigma$ and $\rho$ and two vector of multipliers $\vect{w}$ and $\vect{y}$, obtaining:
\begin{align}
\mathcal{L}_{\sigma,\rho} \left((\tilde{\vect{\gamma}}_\epsilon,\tilde{\vect{z}}),(\vect{\gamma}_\epsilon,\vect{z}),(\vect{y},\vect{w})\right) &= 
\frac{1}{2} \tilde{\vect{\gamma}}_\epsilon^T N \tilde{\vect{\gamma}}_\epsilon + \vect{r}^T_\epsilon \tilde{\vect{\gamma}}_\epsilon  
+  \mathcal{I}_{\tilde{\vect{\gamma}}_\epsilon = \tilde{\vect{z}}}(\tilde{\vect{\gamma}}_\epsilon,\tilde{\vect{z}}) 
+  \mathcal{I}_\Upsilon(\vect{\gamma}_\epsilon) \nonumber \\
+ \vect{w}^T (\tilde{\vect{\gamma}}_\epsilon - \vect{\gamma}_\epsilon) 
&+ \frac{\sigma}{2} \norm{\tilde{\vect{\gamma}}_\epsilon - \vect{\gamma}_\epsilon} 
+ \vect{y}^T (\tilde{\vect{z}} - \vect{z}) 
+ \frac{\rho}{2} \norm{\tilde{\vect{z}} - \vect{z}}
\end{align}
that is also, using the property $\vect{y}^T\vect{r} + \frac{\rho}{2}\norm{\vect{r}} = \frac{\rho}{2}\norm{\vect{r}+\frac{1}{\rho} \vect{y}} - \frac{1}{2\rho} \norm{y}$, the following
\begin{align}
\mathcal{L}_{\sigma,\rho} \left((\tilde{\vect{\gamma}}_\epsilon,\tilde{\vect{z}}),(\vect{\gamma}_\epsilon,\vect{z}),(\vect{y},\vect{w})\right) &= 
\frac{1}{2} \tilde{\vect{\gamma}}_\epsilon^T N \tilde{\vect{\gamma}}_\epsilon + \vect{r}^T_\epsilon \tilde{\vect{\gamma}}_\epsilon  
+  \mathcal{I}_{\tilde{\vect{\gamma}}_\epsilon = \tilde{\vect{z}}}(\tilde{\vect{\gamma}}_\epsilon,\tilde{\vect{z}}) 
+  \mathcal{I}_\Upsilon(\vect{\gamma}_\epsilon) \nonumber \\
+ \frac{\sigma}{2} \norm{\tilde{\vect{\gamma}}_\epsilon - \vect{\gamma}_\epsilon + \frac{1}{\sigma} \vect{w}} 
&- \frac{1}{2\sigma} \norm{w}
+ \frac{\rho}{2} \norm{\tilde{\vect{z}} - \vect{z} + \frac{1}{\rho} \vect{y}}
- \frac{1}{2\rho} \norm{y}
\end{align}

@@

Using these results, and adding a Krasnoselskii-Mann relaxation factor $\alpha \in(0,2]$, one obtains the final algorithm:
%
\begin{subequations}
	\begin{align}
    \begin{bmatrix}
		 H   & D_\epsilon \\
		 D^T & -(\sigma+\rho) I
		\end{bmatrix}
		\begin{Bmatrix}
		 \vect{v}    \\
		 -\tilde{\vect{\gamma}}_\epsilon^{k+1}
		\end{Bmatrix}
		&=
		\begin{Bmatrix}
		 \vect{k} \\
		 -\vect{b}_\epsilon + \sigma \vect{\gamma}_\epsilon^k +\rho \vect{z}^k - \vect{y}^k 
		\end{Bmatrix}
		\\
		\vect{\gamma}^{k+1}_\epsilon &=\tilde{\vect{\gamma}}_\epsilon^{k+1} + (1-\alpha) \vect{\gamma}_\epsilon^k \\
		\vect{z}^{k+1} &= \Pi_\Upsilon \left( \alpha \tilde{\vect{\gamma}}_\epsilon^{k+1} + (1-\alpha) \vect{z}^k  + \frac{1}{\rho} \vect{y}^k \right) \\
		\vect{y}^{k+1} &= \vect{y}^k + \rho \left( \alpha \tilde{\vect{\gamma}}_\epsilon^{k+1} + (1-\alpha) \vect{z}^k - \vect{z}^{k+1} \right) \\
		\vect{r}_{\text{prim}} &= \vect{\gamma}_\epsilon^{k+1}-\vect{z}^{k+1} \\
		\vect{r}_{\text{dual}} &= N \vect{\gamma}_\epsilon^{k+1}+\vect{r}-\vect{y}^{k+1} 
	\end{align}
	\label{eq:admm_final}
\end{subequations}

@@

>>>>

%%
%% This document created by A.Tasora
%%

\documentclass[1p]{elsarticle}

% to typeset URLs, URIs, and DOIs
\usepackage{url}
%\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{cleveref}
\usepackage{bm}
\usepackage{mathtools}
\usepackage{siunitx}
\usepackage{subfig}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{tabularx}
%\usepackage{algorithm}
%\usepackage{algpseudocode}
\usepackage[ruled,vlined, linesnumbered]{algorithm2e}
\usepackage{array}
\usepackage{booktabs}
\usepackage{multicol}
\usepackage{xcolor}
\usepackage[title]{appendix}
\usepackage{empheq}

@@

%% A.TASORA CUSTOM TEX COMMANDS:
%%

\def\avect#1{{\boldsymbol{#1}}}
\def\amatr#1{{\boldsymbol{#1}}}

@@

\begin{document}

%%
%% TITLE
%%

\title{Solving non smooth dynamics using ADMM}

%\author{Alessandro Tasora \\ alessandro.tasora@unipr.it}
\author{Alessandro Tasora}
\ead{alessandro.tasora@unipr.it}
\author{Simone Benatti}
\ead{simone.benatti@studenti.unipr.it}
\author{Dario Mangoni}
\ead{dario.mangoni@unipr.it}
\address{Universit\`a degli Studi di Parma, \\Dpt. of Engineering and Architecture\\Parco Area delle Scienze, 181/A, 43124 Parma, Italy }

% \affil[]{Universit\`a degli Studi di Parma, \\ 
% Dpt. of Engineering and Architecture, Parco Area delle Scienze, 181/A, 43124 Parma, Italy 
% \footnote{E-mail addresses: alessandro.tasora@unipr.it (A. Tasora), \\ simone.benatti@studenti.unipr.it (S. Benatti) dario.mangoni@studenti.unipr.it (D. Mangoni), rinaldo.garziera@unipr.it (R.Garziera).}
% }

\begin{abstract} 
This work discusses an efficient formulation of a geometrically exact three-dimensional beam which can be used in dynamical simulations involving large displacements, collisions and non-linear materials. To this end, we base our model on the shear-flexible Cosserat rod theory and we implement it in the context of Isogeometric Analysis (IGA). 
According to the IGA approach, the centerline of the beam is parameterized using splines; in our work the rotation of the section is parameterized by a spline interpolation of quaternions, and time integration of rotations is performed using the exponential map of quaternions. Aiming at an efficient and robust simulation of contacts, we propose the adoption of a non-smooth dynamics formulation based on differential-variational inequalities. 
The model has been implemented in an open-source physics simulation library that can simulate actuators, finite elements, rigid bodies, constraints, collisions and frictional contacts. 
This beam model has been tested on various benchmarks in order to assess its validity in non-linear static and dynamic analysis; in all cases the model behaved consistently with theoretical results and experimental data. 
\end{abstract}

\begin{keyword}
ADMM \sep non-smooth dynamics \sep contact
\end{keyword}

@@

\begin{abstract}
We propose to use the Alternating Direction Method of Multipliers (ADMM) for solving the variational inequality that arises in non-smooth contact problems in multibody simulations. The ADMM method is based on simple computational primitives and offers good convergence properties especially for scenarios where loose tolerance on the precision can be accepted, thus making it attractive also for real-time applications. The method scales well with the problem size and require a small number of tuning parameters. Differently from other methods known in literature, such as fixed point iterations, it can easily handle problems that mix finite elements and rigid bodies, and it is only marginally affected by odd mass ratios or ill-posed problems. Another attractive feature is that it can be easily warm-started.
\end{abstract}
%
\begin{keyword}
ADMM \sep non-smooth dynamics \sep contact
\end{keyword}

@@

The Alternating Direction Method of Multipliers (ADMM) has been proposed in the 1970s 
\cite{Glowinski75} % [ ADMM invention ]
\cite{Gabay1976} % [ ADMM invention ]
as a practical way to solve constrained optimization problems by iterating over the minimization of two (not necessarily differentiable) objectives.
ADMM introduces auxiliary variables and cast the original problem as the minimization of a separable objective subject to a linear constraint between the original variables and the auxiliary variables. With a proper choice of auxiliary variables, the minimization of the separable function is obtained iterating on two distinct minimization steps of lower complexity. The method iterates updating primal variables, auxiliary variables and dual variables similarly to a fixed point iteration: as such, they compares less favorably to IPMs if one is interested in convergence to high precision, however recent developments showed that in practical scenarios they offer superior speed, scalability and robustness, moreover their structure map well into parallel computing architectures
 \cite{Boyd2011}. % [ ADMM general theory and development ].
These positive properties triggered a recent revival of ADMM and similar operator-splitting methods even as alternatives to IPMs, especially in distributed convex optimization with large-scale problems where moderate precision is acceptable
\cite{Cannon2019}. % [COSMO  conic operator splitting method for large convex problems].

ADMM methods have been used in many fields, for instance computer graphics 
 \cite{Zhang2019}
, computational fluid dynamics
 \cite{Gregson2014} % [ ADMM and fluid simulation, inverse problems ]
, simulation of deformable structures 
 \cite{Overby2017} % [ ADMM projective dynamics hyperelastic models ]
and contact dynamics 
 \cite{Daviet2020} % [ ADMM thin objects frictional contact ]
, to name a few.

Recent developments addressed the possibility of accelerating the ADMM method.
In 
 \cite{Goldfarb2013} % [ ADMM with acceleration, both convex (non-strongly-convex require step skipping), one differentiable]
an acceleration scheme has been proposed for special cases of symmetric ADMMs, and assuming that one of the two objectives is differentiable.
%
In 
 \cite{Goldstein2014} % [ ADMM and AMA with Nesterov. Convergence ok if both problems are strongly convex (non-strongly-convex require restarting), one is quadratic ].
the Nesterov acceleration method has been proposed for problems where at least one of the two objectives is quadratic, and under the assumption that both objectives are strongly convex; if the latter assumption is not guaranteed, restarting strategies are needed. 
%
In 
 \cite{Kadkhodaie2015} %  [ A2DM2 accelerated ADMM, strongly convex ]
a similar acceleration method has been proposed, with $\mathcal{O}(\frac{1}{k^2})$ convergence rate, but without the assumption that one of the objectives must be quadratic.
%
Other types of improvements, based on the Anderson fixed point acceleration, have been put forward in 
 \cite{Zhang2019} % [ ADMM + Anderson acceleration, computergraphics ].
 \cite{Ouyang2020} % [ ADMM + Anderson acceleration, using Douglas-Rachford splitting, computergraphics ]
, showing high performance in computer graphics applications.

Recent developments also focused on the possibility of parallelizing the ADMM iterations on GPU and parallel architectures
\cite{Schubiger2020} % [ GPU implementation of ADMM | no need to compute residuals each iteration | krylov solver each iteration ].
.

In
\cite{Stellato2020} % [OSQP: an operator splitting solver for quadratic programs].
a specialized ADMM method for quadratic problems with conic constraints has been presented: on the top of this method we develop a solver that can exploit the nature of non-smooth dynamical problems. An attractive property of such ADMM is that it requires the solution of a linear system per each iteration, but unlike IPMs, such linear system most often remains unchanged during the iterations, so a factorization can be reused multiple times with a great benefit in terms of speed.

In our embodiment, unknowns are primal variables (velocity measures) and dual variables (impulses in contact points and joints), the convex cone constraints stem from the Coulomb friction law, and the system matrix includes terms from the mass matrices and from the tangent stiffness/damping matrices (hence it is sparse, but not necessarily diagonal).

This paper will present a model for the non-smooth dynamics, it will show how to cast it as a QP with convex conic constraint, it will present the ADMM method to solve it, it will discuss practical implementation details and then it will show benchmarks.

@@

We remark that the ADMM method just makes the assumption of $\Upsilon$ being convex, so the problem can be generalized to $\vect{\gamma}_\epsilon \in \mathcal{C}$ where $\mathcal{C}$ is a generic convex set and at the same time we assume an associated 
 \footnote{Not all dissipative constitutive models are associated, for example in some computational plasticity models the plastic flow might deviate respect to the normal to the yeld surface, and the Coulomb contact model itself is not associated -in fact the contact velocity $\vect{u}$ can range into a wider cone than the polar of the friction cone, but our DVI model compensates fur such effect. We will assume associated models heretofore, and we do not deal with non-associated models to avoid additional complexity for the moment.} 
flow $\vect{u}_\epsilon \in - \mathcal{N}_{C}(\vect{\gamma}_\epsilon)$. For instance, $\mathcal{C}$ could be a capped friction cone to represent plasticization of contacts, or a cone translated downward to represent possible adhesion in contact up to a threshold, or the Von Mises yield region if $\vect{\gamma}_\epsilon$ represent stresses in finite elements undergoing plasticization.

@@

For variables $(\vect{x},\vect{z}) \in \mathbb{R}^m \times \mathbb{R}^n$, the ADMM methods solve constrained optimization problems with separable structure
\begin{subequations}
	\begin{align}
    \text{min} \quad  & f(\vect{x}) + g(\vect{z}) \\
	  \text{s.t.} \quad &  A \vect{x} + B \vect{z} - \vect{b} = \vect{0}
	\end{align}
	\label{eq:admm}
\end{subequations}
by introducing an augmented Lagrangian
\begin{align}
\mathcal{L}_{\rho}(\vect{x},\vect{z},\vect{y})= f(\vect{x}) + g(\vect{z}) + \vect{y}^T (A \vect{x} + B \vect{z} - \vect{b}) + \frac{\rho}{2} \norm{A \vect{x} + B \vect{z} - \vect{b}}
\end{align}
and by iterating over two minimization problems as in the following loop:
\begin{align}
 \vect{x}^{k+1} &\in \text{argmin} \mathcal{L}_{\rho}(\vect{x},\vect{z}^k,\vect{y}^k) \label{eq:lag1}\\
 \vect{z}^{k+1} &\in \text{argmin} \mathcal{L}_{\rho}(\vect{x}^{k+1},\vect{z},\vect{y}^k) \label{eq:lag2}\\
 \vect{y}^{k+1} &= \vect{y}^{k} + \rho ( A \vect{x}^{k+1} + B \vect{z}^{k+1} - \vect{b} )
 \label{eq:admm_iters}
\end{align}
Although the convergence is fast in the first iterations, it tends to deteriorate later as in fixed-point iterations; however in practical scenarios where loose tolerances can be accepted, the method proves to be very efficient and robust.

The performance of ADMM depends on the efficiency of the updates in \eqref{eq:lag1} and \eqref{eq:lag2}. We will design the method such that the step \eqref{eq:lag2} will correspond to a very efficient projection onto a cone with separable structure, whereas the only bottleneck will be \eqref{eq:lag1}, corresponding to a quadratic optimization to be solved with a linear system.

@@

We can write the augmented Lagrangian introducing two step size parameters $\sigma$ and $\rho$ and two vector of multipliers $\vect{w}$ and $\vect{y}$, obtaining:
\begin{align}
\mathcal{L}_{\sigma,\rho} \left((\tilde{\vect{\gamma}}_\epsilon,\tilde{\vect{z}}),(\vect{\gamma}_\epsilon,\vect{z}),(\vect{y},\vect{w})\right) &= 
\frac{1}{2} \tilde{\vect{\gamma}}_\epsilon^T N \tilde{\vect{\gamma}}_\epsilon + \vect{r}^T_\epsilon \tilde{\vect{\gamma}}_\epsilon  
+  \mathcal{I}_{\tilde{\vect{\gamma}}_\epsilon = \tilde{\vect{z}}}(\tilde{\vect{\gamma}}_\epsilon,\tilde{\vect{z}}) 
+  \mathcal{I}_\Upsilon(\vect{\gamma}_\epsilon) \nonumber \\
+ \vect{w}^T (\tilde{\vect{\gamma}}_\epsilon - \vect{\gamma}_\epsilon) 
&+ \frac{\sigma}{2} \norm{\tilde{\vect{\gamma}}_\epsilon - \vect{\gamma}_\epsilon} 
+ \vect{y}^T (\tilde{\vect{z}} - \vect{z}) 
+ \frac{\rho}{2} \norm{\tilde{\vect{z}} - \vect{z}}
\end{align}
that is also, using the property $\vect{y}^T\vect{r} + \frac{\rho}{2}\norm{\vect{r}} = \frac{\rho}{2}\norm{\vect{r}+\frac{1}{\rho} \vect{y}} - \frac{1}{2\rho} \norm{\vect{y}}$, the following
\begin{align}
\mathcal{L}_{\sigma,\rho} \left((\tilde{\vect{\gamma}}_\epsilon,\tilde{\vect{z}}),(\vect{\gamma}_\epsilon,\vect{z}),(\vect{y},\vect{w})\right) &= 
\frac{1}{2} \tilde{\vect{\gamma}}_\epsilon^T N \tilde{\vect{\gamma}}_\epsilon + \vect{r}^T_\epsilon \tilde{\vect{\gamma}}_\epsilon  
+  \mathcal{I}_{\tilde{\vect{\gamma}}_\epsilon = \tilde{\vect{z}}}(\tilde{\vect{\gamma}}_\epsilon,\tilde{\vect{z}}) 
+  \mathcal{I}_\Upsilon(\vect{\gamma}_\epsilon) \nonumber \\
+ \frac{\sigma}{2} \norm{\tilde{\vect{\gamma}}_\epsilon - \vect{\gamma}_\epsilon + \frac{1}{\sigma} \vect{w}} 
&- \frac{1}{2\sigma} \norm{\vect{w}}
+ \frac{\rho}{2} \norm{\tilde{\vect{z}} - \vect{z} + \frac{1}{\rho} \vect{y}}
- \frac{1}{2\rho} \norm{\vect{y}}
\end{align}

@@

Using these results, and adding a Krasnoselskii-Mann relaxation factor $\alpha \in(0,2]$, one obtains the final algorithm:
%
\begin{subequations}
	\begin{align}
    \begin{bmatrix}
		 H   & D_\epsilon \\
		 D^T & -(\sigma+\rho) I
		\end{bmatrix}
		\begin{Bmatrix}
		  \tilde{\vect{v}}^{k+1}  \\
		 -\tilde{\vect{\gamma}}_\epsilon^{k+1}
		\end{Bmatrix}
		&=
		\begin{Bmatrix}
		 \vect{k} \\
		 -\vect{b}_\epsilon + \sigma \vect{\gamma}_\epsilon^k +\rho \vect{z}^k - \vect{y}^k 
		\end{Bmatrix}
		\\
		\vect{\gamma}^{k+1}_\epsilon &=\tilde{\vect{\gamma}}_\epsilon^{k+1} + (1-\alpha) \vect{\gamma}_\epsilon^k \\
		\vect{v}^{k+1} &=\tilde{\vect{v}}^{k+1} + (1-\alpha) \vect{v}^k \\
		\vect{z}^{k+1} &= \Pi_\Upsilon \left( \alpha \tilde{\vect{\gamma}}_\epsilon^{k+1} + (1-\alpha) \vect{z}^k  + \frac{1}{\rho} \vect{y}^k \right) \\
		\vect{y}^{k+1} &= \vect{y}^k + \rho \left( \alpha \tilde{\vect{\gamma}}_\epsilon^{k+1} + (1-\alpha) \vect{z}^k - \vect{z}^{k+1} \right) \\
		\vect{r}_{\text{prim}} &= \vect{\gamma}_\epsilon^{k+1}-\vect{z}^{k+1} \\
		\vect{r}_{\text{dual}} &= N \vect{\gamma}_\epsilon^{k+1}+\vect{r}-\vect{y}^{k+1} 
	\end{align}
	\label{eq:admm_final}
\end{subequations}

@@

\section{Optimizations}

An interesting property of this formulation is that one does not need to recover $\vect{v} = \amatr{H}^{-1}(\vect{k}+\amatr{D}^T \vect)\vect{\gamma}^{k+1}$ at the end because, for $\alpha=1$, this is also a byproduct of the linear system solve as $\vect{v}= \tilde{\vect{v}}^{k+1}$.
In case of $\alpha\neq1$, one needs to keep it updated as 
$\vect{v}^{k+1} =\tilde{\vect{v}}^{k+1} + (1-\alpha) \vect{v}^k$; 
but this also means that an initial value of $\vect{v}^0$, consistent with the initial value $\vect{\gamma}^0$, must be computed, hence before the iteration one has to initialize 
\[
\vect{v}^0 = \amatr{H}^{-1}(\vect{k} + \amatr{D} \vect{\gamma}^0)
\]

We see that the dual residual $\vect{r}_{\text{dual}}$ is computed using the $\amatr{N}$ matrix, but can be inefficient as we do not want to compute it explicitly. One option is to remember the definition of the Schur complement and use 
\[
\vect{r}_{\text{dual}} = \amatr{D}^T \vect{v}^{k+1} - \amatr{E} \vect{\gamma}^{k+1} + \vect{b}_\epsilon + \vect{y}
\]
Another even faster option (but only for $\alpha=1$) would be to compute the dual residual using
\[
\vect{r}_{\text{dual}} = \rho (\vect{z}^{k+1} - \vect{z}^{k} )
\]

TO COMPLETE

\section{Step size selection}

In general, small values of $\rho$ cause a fast reduction of the dual residual, but a slow reduction of the primal residual. This has a nice physical interpretation: truncating iterations based on too small $\rho$ most often lead to solutions where parts do not interpenetrate, but some contact forces do not satisfy the inclusion in the Coulomb cone. That is, some parts might stick like contacts were glued.

On the other hand, large values of $\rho$ cause a fast reduction of the primal residual, but a slow reduction of the dual residual. This means that truncating iterations with too large $\rho$ most often lead to solutions where contact forces satisfy very well the inclusion in the Coulomb friction cones, but parts might interpenetrate.

Clearly, this calls for the introduction of a scheme that adaptively changes the step size $\rho$ in order to achieve a balanced reduction of both residuals.

TO DO

Step size:

In \cite{Wohlberg2017} [ a step size selection via residual balancing, scale residuals given problem measuring units]

In \cite{Xu_adaptive_2017} [ Spectral stepsize selection as in Barzilai Borwein ]

\section{Benchmarks}

TO DO
